---
title: "tooth-growth-analysis"
author: "Mephistopheles-0"
format: pdf
editor: visual
---

# A Basic Statistical Analysis of the ToothGrowth Dataset (The Effect of Vitamin C on Tooth Growth in Guinea Pigs)

## Overview

We're going to analyze the ToothGrowth data in the R datasets package by:

-   Load the ToothGrowth data and perform some basic exploratory data analyses

-   Provide a basic summary of the data.

-   Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

-   State your conclusions and the assumptions needed for your conclusions.

## About the data

The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as `VC`).

## Loading the dataset and the ggplot library

```{r}
# Loading the ToothGrowth dataset
data(ToothGrowth)
library(ggplot2)
```

## Basic infos

```{r}
# Printing the structure of the dataset
str(ToothGrowth)
```

```{r}
# summary for the Supplement type
summary(ToothGrowth$supp)
```

```{r}
# summary for the Dose
summary(ToothGrowth$dose)
```

```{r}
# Displaying the first few rows of the dataset
head(ToothGrowth)
```

```{r}
# Basic statistics summary
summary(ToothGrowth)
```

## Some DataViz

```{r}
# Boxplot for the variable 'len'
boxplot(ToothGrowth$len, 
        main="Boxplot of Length", 
        xlab="Supplement", 
        ylab="Length", 
        col=c("orange","green","blue"), 
        border="brown")
```

```{r}
hist(ToothGrowth$len, 
     main="Histogram of Length", 
     xlab="Length", 
     ylab="Frequency", 
     col="skyblue")
```

```{r}
plot(density(ToothGrowth$len), 
     main="Density Plot of Length", 
     xlab="Length", 
     ylab="Density", 
     col="red")
```

```{r}
# Q-Q plot for checking normality of the variable 'len'
qqnorm(ToothGrowth$len, 
       main="Q-Q Plot of Length")
qqline(ToothGrowth$len)
```

```{r}
cor(ToothGrowth[sapply(ToothGrowth, is.numeric)])
```

We can see that the length of the tooth and the dose of vitamin C have a high correlation.

```{r}
plot1 <- ggplot(ToothGrowth, aes(x = factor(dose), y = len, fill = factor(dose)))
plot1 + geom_boxplot() + xlab("dosage") + ylab("length")
```

again, the boxplot clearly shows that as the dosage of vitamin C increases the length increases also.

## Compare Tooth Growth by Dose and Supplement Type using Confidence Intervals and Testing.

### The 95% Confidence Interval of Tooth Growth

```{r}
# The mean of the tooth length
mu <- mean(ToothGrowth$len)

# The standard deviation of the tooth length
sigma <- sd(ToothGrowth$len)

# the 95% confidence interval
interval <- mu + c(-1,1) * qnorm(0.975) * sigma / sqrt(length(ToothGrowth))
interval
```

### The T-Test by dose and supp

```{r}
t.test(len ~ supp, data = ToothGrowth)
```

thus, we can't reject the null hypothesis .

### T-Test of the Dose level.

```{r}
# create subsets of small, medium, and large dosage
dose1 <- subset(ToothGrowth, dose == 0.5)
dose2 <- subset(ToothGrowth, dose == 1)
dose3 <- subset(ToothGrowth, dose == 2)
```

#### T-Test for small dasage

```{r}
t.test(len~supp, data = dose1)
```

####  T-Test for medium dosage

```{r}
t.test(len~supp, data = dose2)
```

#### T-Test for large dosage

```{r}
t.test(len~supp, data = dose3)
```

-   for the 1st and the 2nd tests we have quietly small p-value, so we can reject the null hypothesis.

-   for the 3rd test, we have a high p-value of 0. 9639, thus we cannot reject the null hypothesis.

## Conclusion

-   The tooth length growth has nothing to do with the supplements.

-   for the 0.5 and 1 dosages, Orange Juice (OJ) has a higher effect on the tooth length comparing the Vitamin C (VC).

-   for the dosage 2, there is no such difference between the OJ and VC supplements effect.
